### 2019-03-23 paper.tex
fleshed out middle sections

\documentclass{article}
\title{The Epistemic Modal \emph{probable}}
\author{Conrad Friedrich \\ MCMP, LMU Munich }
\date{\today}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{enumitem}
\usepackage{setspace}

\usepackage{ebproof/ebproof}
\usepackage{calrsfs}
% Used for todo notes
\usepackage{xcolor}

\usepackage[backend=biber,authordate,
ibidtracker=context,natbib,doi=false,isbn=false,url=false]{biblatex-chicago}

\addbibresource{/home/gilbert/Documents/bibliography/references.bib}
\theoremstyle{definition}

\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}

\newtheorem{definition}{Definition}

\newcommand\todo[1]{\textcolor{red}{#1}}

\renewcommand{\L}{\pazocal{L}}
\newcommand{\F}{\pazocal{F}}
\newcommand{\lb}{\llbracket}
\newcommand{\rb}{\rrbracket}
\newcommand{\Po}{\text{Po}}

\newcommand{\succeqK}{\succeq^K_{M,w}}
\begin{document}

\onehalfspacing

\maketitle

\tableofcontents
\section{Introduction}

\todo{
  \begin{itemize}[nosep]
    \item Write an Introduction
     \item be perfectly clear as to what happens. SIGNPOST 
  \end{itemize}
}
\section{Intensional Semantics}
\todo{ \begin{itemize}[nosep]
  \item Tie the paper into framework used in class
\end{itemize} }

Note that we don't need first order logic. 
Abstracted away from syntax, just look at the semantics.
NEAT summary in Ivano's script, Class 9, page 3/4
\section{The Epistemic Modal \emph{probable}}
\todo{\begin{itemize}[nosep]
    \item  Give some natural language motivation
    \item Also show the need for expanding the standard intensional semantics!
    \item How is probable to be evaluated? Epistemic interpretation? MacFarlane's Solipcism.
    \item Mention how Lassiter has a different, graded adjective based approach.
    \item Discuss how absolute probability is best reduced to relative likelihood with the negation
      \item note that there are a lot more interesting problems associated with probable but to have an readable paper need to focus on particulars.
  \end{itemize}
}

To base the following discussions on a common language, we define a
propositional language $\L$ like \textcite{harrison-trainor17_prefer} and
\textcite{holliday13_measur} do:
\[
\varphi ::= p ~|~ \neg \varphi ~|~ (\varphi \land \psi) ~|~  (\varphi \succeq \psi) ~|~ \triangle\varphi ~|~ \Diamond \varphi
\]
Where $p \in \text{Atom}$, i.e. an atomic formula. The other boolean connectives $(\varphi \rightarrow \psi), (\varphi \vee
\psi)$ as well as $\top, \bot$ are defined in the standard way from negation and conjunction.
Define further $\Box \varphi = \neg\Diamond\neg \varphi$.
The intended readings are as follows:
\begin{align*}
  \varphi \succeq \psi &: \varphi \text{ is at least as likely as } \psi. \\
  \triangle \varphi &:\text{probably } \varphi. \\ 
                      \Diamond \varphi &: \text{It might be that } \varphi.
\end{align*}

\subsection{Criteria for Adequacy}
In his \parencite*{yalcin10_probab_operat}, Yalcin critically evaluates Kratzer's relative likelihood account and also Hamblin's account, which is based on the notion of possiblity (resp. plausiblity) functions \parencite{hamblin59_modal_probab}.
Yalcin does this by following the good philosophical practice of establishing a set of desiderata the accounts can be evaluated against. In the case for judging semantics for a natural language phenomenon, these naturally come in the form of inference patterns that strike one as intuitively valid or invalid.
The practice is fruitful because it gives criteria to compare different accounts in a transparent and comprehensible fashion, and shifts the focus of the problematic notion of \emph{intuitive plausibility} from judging the accounts themselves to judging the plausibility of the inference patterns.
Whether some formally specified semantics validates a given inference pattern is then simply not a matter of debate and can be shown mathematically given the assumptions one is forced to make explicit.   
\todo{
\begin{itemize}[nosep]
  \item Motivate and describe the conditions
\end{itemize}
}
\begin{center}
    \begin{prooftree}
        \hypo{ \triangle \varphi } \infer1[(V1)]{ \neg \triangle \neg \varphi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
        \hypo{ \triangle (\varphi \land \psi)} \infer1[(V2)]{ \triangle \varphi \land \triangle \psi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
        \hypo{ \triangle \varphi} \infer1[(V3)]{ \triangle (\varphi \lor \psi)}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ } \infer1[(V4)]{ \varphi \succeq \bot}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ } \infer1[(V5)]{ \top \succeq \varphi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ \Box \varphi } \infer1[(V6)]{ \triangle \varphi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ \triangle \varphi } \infer1[(V7)]{ \Diamond \varphi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ \psi \succeq \varphi } \hypo{\triangle \varphi} \infer2[(V11)]{ \triangle \psi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ \psi \succeq \varphi } \hypo{\varphi \succeq \neg\varphi}\infer2[(V12)]{ \psi \succeq \neg\psi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ \varphi \succeq \psi } \hypo{\varphi \succeq \chi}\infer2[(I1)]{ \varphi \succeq (\psi \lor \chi)}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ \varphi \succeq \neg \varphi } \infer1[(I2)]{ \varphi \succeq \psi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ \triangle \varphi } \infer1[(I3)]{ \varphi \succeq \psi}
    \end{prooftree}
\end{center}

\subsection{Criteria not Considered}
Yalcin and also \textcite{lassiter10_gradab} introduce an additional criterion motivated by examples of the following sort:
\begin{enumerate}[nosep]
  \item There is a 60\% probability that it is raining.
  \item $\varphi$ is twice as likely as $\psi$.
\end{enumerate}
Yalcin argues that sentences of this type are ``used and interpreted by speakers who are innocent of the mathematical theory of probability'' \parencite[][p.~923]{yalcin10_probab_operat} and that a semantics for probability operators in natural language must be able to account for them, while Lassiter merely states it as a problem for preferential semantics \parencite[][p.~210]{lassiter10_gradab}. Both share the implicit assumption that both the qualitative talk (`probably $\varphi$') and the quantitative talk like in (2) above need to be accounted for by the same semantics. Given the methodological practice of frequent appeal to intuition to support claims of this sort, the assumption can also be doubted on the same grounds. It is equally plausible that people usually employ different notions of \emph{probable} (or \emph{likely}) in the qualitative and the quantitative case. Qualitative talk is quotidian and ubiquitous, whereas quantitative talk implies the notion of a probability function. It might be the case that speakers use it `innocently' without recourse to probability theory, but it's a different question of whether they do so competently. We argue that they might as well don't. The result is that the criterion does not necessarily help us evaluate adequate semantics for likelihood talk, and thus will be excluded from the present considerations.

\section{Kratzer's Preferential Semantics}
\todo{
  \begin{itemize}
    \item Clean up this section a bit
  \end{itemize}
}
For \textcite{kratzer91_modal} epistemic modals like \emph{probably} are just a special case of a more general class of modal auxiliaries.
These can be epistemic (e.g. `Jockl must have been the murderer [in view of the available evidence]'), deontic (e.g. `Jockl must go to jail [in view of what the law provides]'), or circumstancial (e.g. `Jockl can lift the rock [given the weight of the rock and the conditions of Jockl's muscles, etc.]').
Further, there are teleological (in view of certain aims), bouletic (in view of certain wishes), and doxastic (in view of certain beliefs) modalities.
This list is by no means complete.

Kratzer postulates that modality is always relative to a certain kind of interpreting the modalities involved, following Peirce (cite).
She set out to formulate these different modalities in a common framework such that we'd only need to specify some parameters of the framework to determine the relevant modalities and give truth conditions for sentences involving modalities, while focussing on epistemic modals.

Kratzer is unhappy with what she calls the \emph{standard analysis} as it gives the wrong verdict in two salient scenarios.
These are concerned with conflicting rules of law and an a combination of indicative conditionals and modals, respectively.
We will not reiterate these scenarios here, but--spoiler alert--Kratzer's theory manages to do them justice.

More relevantly, the account for modality according to Kratzer needs to include graded modalities, of the following sort:
\begin{enumerate}[nosep,label=\alph*)]
  \item Michl must be the murderer.
  \item Michl is probably the murderer.
  \item There is a good possibility that Michl is the murderer.
  \item Michl might be the murderer.
\end{enumerate}
Here, the modalities are listed in decreasing strength, such that the intended reading is that a) implies b), b) implies c), and c) implies d).
Note that is a qualitative approach to graded modals, as opposed to quantitative approaches, employing a scale structure in the semantics, as described by \textcite{lassiter10_gradab}.

\todo{\begin{itemize}[nosep]
  \item The account presented has a couple of advantages.
\item Neatly fits with the other modals, just change f and g
\item Gives an account of probable without resorting to probability functions (give argument by Hamblin - maybe more in philosophy texts? check titelbaum)
\end{itemie}
}
How do the semantics look like?
We assume a set $W$ of possible worlds, a set of propositions $P$ such that elements of $P$ are subsets of $W$, and assume that a proposition $p$ is true at $w$ iff $w \in p$.
The crucial element in Kratzer's semantics that distinguishes her account from most simple semantics via Kripke frames with an accessibility relation is the \emph{conversational background}.
It has two components.

The \emph{modal base} determines for every world the set of worlds which are epistemically accessible from it.
More formally, the modal base is a function $f: W \rightarrow \wp(W)$ assigning each world $w$ a set of worlds.
The set of worlds $f(w)$ assigned to $w$ represents, in the epistemic case, what we know at $w$.

The \emph{ordering source} is a function $g: W \rightarrow \wp(W)$ s.t. each $w$ is assigned a set of propositions $g(w)$ which interpreted as an \emph{ideal} (cf. \cite{lewis81_order_seman_premis_seman_count}). The ordering source induces an order $\leq_{g(w)}$ on $W$ for all $w,w' \in W$ as follows:
\begin{equation}
    \label{eq:porder}
w \leq_{g(w)} w' :\iff \{p: p\in g(w) ~\&~ w' \in p\} \subseteq \{p: p\in  g(w) ~\&~ w \in p\}
\end{equation}
\todo{THIS NEED UPDATING SINCE g(w) CHANGED!}
\begin{equation}
    \label{eq:porder}
w \leq_{g(w)} w' :\iff \{A: g(w) \subseteq A ~\&~ w' \in A\} \subseteq \{A: g(w) \subseteq A ~\&~ w \in A\}
\end{equation}
The intended reading is that $w$ is at least as close to the ideal represented by $g(w)$ as $w'$ iff all the ideal propositions in $g(w)$ which are true in $w'$ are also true in $w$.
From this, Kratzer defines a the semantics for a list of different expressions, like `good possibility', `possibility' and so on.
For our purposes, it suffices to look at those definitions involved in defining the meaning of \emph{probably $\alpha$}.

First, define `as least as good a possibility as'.
The induced order on worlds $\leq_{g(w)}$ is `lifted' to an order on the level of propositions.
For a modal base $f$ and a ordering source $g$, for propositions $A,B$ and world $w$, we can define
\begin{equation}
    \label{eq:lifted}
    A \succeq_w B :\iff \forall b \in B_w \exists a \in A_w: a \leq_{g(w)} b,
\end{equation}
where $A_w = f(w) \cap A$, and $B_w$ analogously (cf. \cite[][p.~519]{holliday13_measur}).
In words, a proposition $A$ is as least as good a possibility as a proposition $B$ seen from a world $w$ iff for any world in $B$ epistemically accessible from $w$ there is a world in $A$ accessible from $w$ that is at least as close to the ideal picked out by $g(w)$.

The strict part is defined as one would expect:

\noindent $A \succ_w B :\iff {A \succeq_w B ~\&~ B \not\succeq_w A}$.

Finally, we can assign a meaning to \emph{probably $\alpha$} given a set of worlds $W$, a modal base $f$ and a ordering source $g$ \parencite[][p.~645]{kratzer91_modal}:
\begin{equation}
\llbracket \text{probably } \alpha \rrbracket^{f,g,w} = 1 :\iff \lb \alpha \rb \succ_w \lb \neg\alpha \rb
\end{equation}
Assuming here that we are looking at a flat fragment and $\alpha$ doesn't contain modals itself.
Intuitively,... [briefly explain]

\subsection{Formal Setup}

To put it slightly more formally, we define: 
\begin{definition}
    A \emph{Kratzer preferential model} is a tuple $M = \langle W,f,g,V \rangle$ s.t.
    \begin{itemize}[nosep]
\renewcommand\labelitemi{--}
      \item $W$ is a non-empty set of possible worlds,
      \item $f,g$ are functions $f,g: W \rightarrow \wp(W)$,
      \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$
    \end{itemize}
\end{definition}

Further, we want the induced order defined above also hold for $M$.
This is done in the following uninteresting definition:
Let $f$ be a modal base, $g$ be a ordering source, $W$ a set of worlds.
For each $w \in W$ let $\succeq_w$ be the order on subsets of $W$ defined in
Equation \ref{eq:lifted}.
Let $M = \langle W,f,g,V \rangle$ be a Kratzer preferential model.
For $p,q \subseteq W$ we define an induced order on $\wp(W)$:   
\begin{equation}
    \label{eq:induced}
       p \succeq^{M,w} q \text{ iff } p \succeq_w q.
\end{equation}

\begin{definition}[Kratzer preferential semantics]
    \label{def:kratsem}
    Let $M = \langle M,f,g,V \rangle$ be a Kratzer preferential model, $w \in W$, $\varphi, \psi \in \L$ and let $\succeq^{M,w}$ be the induced order on $\wp(W)$ by Equation \ref{eq:induced}. We define

    \begin{enumerate}[nosep]
      \item If $p \in \text{Atom}$, $M,w \models p$ iff $w \in V(p)$,
      \item $M,w \models \neg \varphi$ iff $M,w \not\models \varphi$
      \item $M,w \models (\varphi \land \psi)$ iff $M,w \models \varphi$ and $M,w \models \psi$,
      \item $\lb \varphi \rb^M = \{w' \in W : M,w' \models \varphi\}$,
      \item $M,w \models \varphi \succeq \psi$ iff $\lb \varphi \rb^M \succeq^{M,w} \lb \psi \rb^M$,
      \item $M,w \models \triangle \varphi$ iff $M,w \models \varphi \succeq \neg \varphi ~\&~ \neg (\neg \varphi \succeq \varphi)$,
      \item $M,w \models \Diamond \varphi$ iff ...
    \end{enumerate}
\end{definition}

\todo{FORMALIZE MIGHT PHI}
\subsection{Discussion}

How does Kratzer's account fare with respect to the adequacy conditions we defined?
The semantics validate conditions (V1-V7), which can be easily seen by applying the definitions.
For example, let us establish that Kratzer preferential semantics satisfy (V2), i.e.: 
\begin{center}
\begin{prooftree}
        \hypo{ \triangle (\varphi \land \psi)} \infer1[(V2)]{ \triangle \varphi \land \triangle \psi}
    \end{prooftree}
\end{center}
Suppose $M,w \models \triangle (\varphi \land \psi)$.
Then $\lb (\varphi \land \psi) \rb^M \succ^{M,w} \lb \neg(\varphi \land \psi)\rb^M$.
Let $y \in \lb \neg \varphi \rb^M$.
Then $y \in \lb \neg \varphi \lor \neg\psi \rb^M = \lb \neg(\varphi \land \psi)\rb^M$.
By assumption, there is an $x \in \lb (\varphi \land \psi) \rb^M$ such that $x \leq_{g(w)} y$.
But then $x \in \lb \varphi \rb^M$, too.
Hence $\lb \varphi \rb^M \succ^{M,w} \lb \neg \varphi \rb^M$ and thus $M,w \models \triangle \varphi$.
The case for $M,w \models \triangle \psi$ works analogously.

Kratzer's account does \emph{not} validate (V11-V12).
\textcite[pp.~935, note~8]{yalcin10_probab_operat} gives a countermodel. In addition, the account \emph{does} validate pattern (I1), as can easily be seen. This is already an unwelcome result. Additionally, (I1) together with (V5) entails (I2) \parencite[p.~922]{yalcin10_probab_operat}. That is, on this account, anything that is probable is a least as likely as any tautology. This is, of course, a ruinous result.

\section{Hamblin's Possibility Semantics}
Already in the 1950's, \textcite{hamblin59_modal_probab} developed a semantics for the \emph{probable}-operator based on what he called plausibility and more recently is referred to as possiblity. Possibility functions are quite similar to probability functions, but do not require finite additivity. They have been used to model an agent's belief state when the focus is on modeling the agent's surprise as well as ignorance towards some proposition. They are useful in modeling vagueness and have the advantage over probability functions that the possibility of the union of two propositions is uniquely determined by the marginal possibilities of the propositions, while probabilities only provide bounds unless the propositions are disjoint \parencite[][p.~45]{halpern03_reason_about_uncer}.

Why should possibilities be particularly appropriate in semantics? Hamblin does not present a clear argument, but refers to that the weighing of reasons is captured better in possibilities than in probabilities \parencite[][p.~240]{hamblin59_modal_probab}.
Yalcin gives his interpretation of Hamblin as that probabilities are too idealized, and possibilities provide a less restricted formal structure and are thus more appropriate to model everyday likelihood talk.
In any case, let us look at the formal setup and evaluate whether Hamblin's account holds up to our criteria. 
The account presented here is a reinterpretation based on \textcite{hamblin59_modal_probab}, \textcite[][pp.~42]{halpern03_reason_about_uncer} and \textcite{yalcin10_probab_operat} to fit the account into contemporary semantics.

\subsection{Formal Setup}
\noindent Let $W$ be a set. A boolean algebra on $W$ is a set $\F \subseteq \wp(W)$ such that
\begin{enumerate}[nosep]
  \item $\emptyset, W \in \F$,
  \item $A \in \F \Rightarrow \bar{A} \in \F$, and
  \item $A,B \in \F \Rightarrow A \cup B \in \F$.
\end{enumerate}

\noindent Let $\F$ be a boolean algebra. A possibility function is a function $p: \F \rightarrow [0,1]$ such that
\begin{enumerate}[nosep]
  \item $p(\emptyset) = 0$,
  \item $p(W) = 1$, and
  \item $p(U \cup V) = \max(p(U),p(V))$.
\end{enumerate}
\begin{definition}
    A \emph{Hamblin possibility model} is a tuple $M = \langle W,R,V,\mathcal{F},\Po\rangle$ such that
    \begin{itemize}[nosep]
        \renewcommand\labelitemi{--}
      \item $W$ is a set of possible worlds,
      \item $R$ is a binary relation on $W$,
        \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$,
        \item $\mathcal{F}$ is a function assigning each world $w \in W$ a boolean algebra $\F_w$ on $R(w)$, and 
          \item $\Po$ is a function assigning each world $w \in W$ a possibility function \\ $p_w: \F_w \rightarrow [0,1]$.
    \end{itemize}
\end{definition}

\begin{definition}[Hamblin possiblity semantics]
    \label{def:hambsem}
    Let $M = \langle M,R,V,\mathcal{F},\Pr \rangle$ be a Hamblin possiblity model, $w \in W$, $\varphi, \psi \in \L$. We define clauses 1-4 just like in Kratzer preferential semantics (Definition \ref{def:kratsem}). We define further:
    \begin{enumerate}[nosep]
  \setcounter{enumi}{5}
      \item $M,w \models \varphi \succeq \psi$ iff $p_w(\lb \varphi \rb^M) \geq p_w(\lb \psi \rb^M) $
      \item $M,w \models \triangle \varphi$ iff $M,w \models \varphi \succ \neg \varphi$ 
    \end{enumerate}
\end{definition}
\subsection{Discussion}
\todo{
  \begin{itemize}
    \item Write Discussion
  \end{itemize}
}
\section{Yalcin's Probability Semantics}

\todo{
  \begin{itemize}
    \item Write Introduction
  \end{itemize}
}
\subsection{Formal Setup}

\noindent Let $\F$ be a boolean algebra. A probability function is a function $\mu: \F \rightarrow [0,1]$ such that
\begin{enumerate}[nosep]
  \item $\mu(\emptyset) = 0$ 
  \item $\mu(W) = 1$
  \item $\mu(U \cup V) = \mu(U) + \mu(V) \text{ if } U \cup V = \emptyset $
\end{enumerate}

\begin{definition}
    A \emph{Yalcin probability model} is a tuple $M = \langle W,R,V,\mathcal{F},\Pr\rangle$ such that
    \begin{itemize}[nosep]
        \renewcommand\labelitemi{--}
      \item $W$ is a set of possible worlds,
      \item $R$ is a binary relation on $W$,
        \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$,
        \item $\mathcal{F}$ is a function assigning each world $w \in W$ a boolean algebra $\F_w$ on $R(w)$, and 
          \item $\Pr$ is a function assigning each world $w \in W$ a probability function \\ $\mu_w: \F_w \rightarrow [0,1]$.
    \end{itemize}
\end{definition}

\begin{definition}[Yalcin probability semantics]
    \label{def:yalcsem}
    Let $M = \langle M,R,V,\mathcal{F},\Pr \rangle$ be a Yalcin probability model, $w \in W$, $\varphi, \psi \in \L$. We define clauses 1-3 just like in Kratzer preferential models (Definition \ref{def:kratsem}). We define further:
    \begin{enumerate}[nosep]
  \setcounter{enumi}{4}
      \item $M,w \models \varphi \succeq \psi$ iff $\mu_w(\lb \varphi \rb^M) \geq \mu_w(\lb \psi \rb^M) $
      \item $M,w \models \triangle \varphi$ iff $M,w \models \varphi \succ^{M,w} \neg \varphi$ iff $ \mu_w(\lb \varphi \rb^M) > .5$
    \end{enumerate}
\end{definition}

\subsection{Discussion}
\todo{
  \begin{itemize}
    \item Write Discussion
  \end{itemize}
}
\section{Holliday and Icard's Preferential Semantics}
\todo{
  \begin{itemize}
    \item Write Introduction
  \end{itemize}
}
\subsection{Formal Setup}
\begin{definition}
    A Holliday and Icard preferential model is a tuple $M = {\langle
      W,R,V,\succeq\rangle}$ such that
    \begin{itemize}[nosep]
        \renewcommand\labelitemi{--}
      \item $W$ is a set of possible worlds,
      \item $R$ is a binary relation on $W$,
      \item $V$ is a function $V: \text{ Atom } \rightarrow \wp(W)$, and
      \item $\succeq$ is a function assigning each world $w \in W$ a preorder
        (reflexive and transitive binary relation) $\geq^{M,w}$ on $W$. 
    \end{itemize}
\end{definition}
Let $R[w] = \{w':wRw'\}$. For any proposition $P \subseteq W$, $P_w := P \cap R[w]$. Further, for two proposition $A,B \subseteq W$ we define
\begin{multline}
    \label{eq:injection}A \succeq^{M,w} B :\iff \text{ there exists an injective function } f: B_w \rightarrow A_w \\ \text{ such that } f(x) \geq^{M,w} x \text{ for all } x \in B_w.
\end{multline}

\begin{definition}[Holliday and Icard preferential semantics] Let $M = {\langle
      W,R,V,\succeq \rangle}$, $w \in W, \varphi,\psi \in \L$ and let $\succeq^{M,w}$ be the order on $W$ defined by Equation \ref{eq:injection}. We the define clauses 1-6 just like in Kratzer preferential semantics (Definition \ref{def:kratsem}).
\end{definition}

Holliday and Icard's semantics is very similar to Kratzer's.
In particular, we can associate the modal base and ordering source of Kratzer's with the binary relation $R$ and the $\geq$-function.
Just as the modal base determines for each world a set of epistemically accessible worlds, so does $R[w]$.
And just as the ordering source induces an order on propositions, so does $\geq$.
For our present purposes, both formulations are equivalent even.
Holliday and Icard make a crucial change to how the order is induced, however.
To see this, note that Equation \ref{eq:lifted} can be rewritten as an equivalent definition:

For a modal base $f$ and a ordering source $g$, for propositions $A,B$ and world $w$, define
\begin{multline}
    \tag{2*}
    \label{eq:lifted-star}
    A \succeq_w B :\iff \text{ there exists a function } f: B_w \rightarrow A_w \\ \text{ such that } f(x) \leq_{g(w)} x \text{ for all } x \in B_w.
\end{multline}
Compared to Equation \ref{eq:injection}, besides the inverted direction of $\geq$ which is just due to how the notions were defined, the difference is that Holliday and Icard require an injective function assigning each element of $B$ a unique element of $A$ such that the element of $A$ is more plausible, normal etc. This simple change has interesting consequences, as we will see in the next section. 
\subsection{Discussion}
\todo{
  \begin{itemize}
    \item Write Discussion
  \end{itemize}
}
\section{Conclusion}
\todo{
  \begin{itemize}
    \item Write Conclusion and further studies
  \end{itemize}
}
What about the $\epsilon$-semantics (Adams, Pearl)? 
\nocite{hamblin59_modal_probab,holliday13_measur,harrison-trainor17_prefer,kratzer91_modal,lassiter10_gradab,yalcin10_probab_operat,kratzer98_seman}
 \printbibliography
\end{document}